---
title: "Pip"
output: html_document
---
#Install packages ----

```{r setup, include=FALSE}
install.packages("car")
install.packages("agricolae") 
rm(list=ls()) #Clear the global environment (whenever you start a new analysis) ----
```

#Import your data ----
#(import an excel spreadsheet saved in csv format) 

```{r cars}
setwd ("/Users/kajeethasarvananthan/Desktop") #set the working directory to wherever your file is located 
mydata <- read.csv("mydata.csv", header=TRUE)
```

#Set the variables as factors or numeric ----

```{r pressure, echo=FALSE}
mydata$Treatment <- factor(mydata$Treatment)
mydata$Plant <- factor(mydata$Plant)
mydata$cfu <- as.numeric(mydata$cfu)
```
#ask for whatever type of basic output you'd like, for example... 
```{r}
print(mydata) 
plot(mydata$Treatment,mydata$cfu) #can be useful for spotting outliers 
plot(mydata$Plant,mydata$cfu)
```
#you can check the mean, standard deviation, number of entries etc
```{r}
tapply(mydata$cfu,mydata$Treatment,mean)
tapply(mydata$cfu,mydata$Plant,mean)
tapply(mydata$cfu,mydata$Treatment,sd)
tapply(mydata$cfu,mydata$Plant,sd)
tapply(mydata$cfu,mydata$Treatment,length)
tapply(mydata$cfu,mydata$Plant,length)
```
#Run ANOVA ---- 
#for 2 factor interaction, use colon VS for separate effects, use star
```{r}
my_anova_model = aov(cfu ~ Treatment:Plant, data=mydata) #the numeric variable must be first in this expression 
summary(my_anova_model)#show results (NOTE: before interpreting the results you need to check that your data meet the assumptions of the model [if not, the results could be nonsense]) 
```

#ANOVA results significant if F<0.05

#Test assumptions of the ANOVA model ----

```{r}
shapiro.test(mydata$cfu) #test for normality of the data (numeric variable), NOT IMPORTANT?
                         #null hypothesis: data = normally distributed, thus should get p>0.05

hist(mydata$cfu) #visual assessment of normality 

shapiro.test(resid(my_anova_model)) #test for normality of the residuals, IMPORTANT
                                #null hypothesis = normally distributed, thus should get p>0.05
hist(resid(my_anova_model))

plot(my_anova_model) #the first plot can be used to assess homoscedasticity of the residuals; you're hoping for a roughly equal distribution about zero (no fan shapes) 

bartlett.test(cfu ~ Plant:Treatment, data=mydata) #ONLY 1-WAY ANOVA. test for homegeneity of variance
                                                  #null hypothesis: homogeneity of variance
                                                  #apparently test is sensitive to departures from normality and 
```
#Levene's test is better at dealing with this

```{r}
library(car) 
leveneTest(cfu ~ Plant:Treatment, data=mydata) #test for homogeneity of variance 
                              #null hypothesis: homogeneity of variance, thus should get p>0.05
```
#Post-hoc Tukey HSD test ----
```{r}
library(agricolae)                                                       
TukeyHSD(my_anova_model) #gives p values but have to figure out the letter codes for yourself 
print(HSD.test(my_anova_model, "mydata$treatment")) #ONLY 1-WAY ANOVA. gives letter codes but no p values
options(max.print = 9000) #execute this if rows in output get ommitted due to long length, then re-execute TukeyHSD
```
#If tests were failed, log transform, then redo ANOVA and Tukey HSD ----
#if your data fail to meet any of the assumptions, log transform it and create another ANOVA model using this new variable 
```{r}
mydata$logcfu = log10(mydata$cfu)
mydata$logcfu <- as.numeric(mydata$logcfu)
my_anova_model_2 = aov(logcfu ~ Treatment:Plant, data=mydata) #for 2 factor interaction, use colon VS for separate effects, use star
summary(my_anova_model_2)
```
#test assumptions of the model 
```{r}
shapiro.test(mydata$logcfu)  
hist(mydata$logcfu)
shapiro.test(resid(my_anova_model_2))
hist(resid(my_anova_model_2))
plot(my_anova_model_2)

leveneTest (logcfu ~ Treatment*Plant, data=mydata)
```
#if your data satisfy the assumptions & ANOVA returned a significant result, proceed to post-hoc testing
```{r}
library(agricolae)
TukeyHSD(my_anova_model_2) #gives p values but have to figure out the letter codes for yourself 
print(HSD.test(my_anova_model_2, "mydata$treatment")) #gives letter codes but no p values (only for one-way ANOVA)
```


